### 2015-01-26 ASP.tex
update

This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving. 
\annotMF{Source ? Sinon : “one of the most” $\Rightarrow$ “very”} \repEB{OK}

@@

\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
\repEB{T’as raison au début j’ai écrit tout le code en mode $math$ et puis j’ai changé juste la partie du code et j’ai oublié de faire la mise à jour de cette partie de définition d’ASP}
The other classes remain constant as they meet some informal agreement. In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).

@@

\item A \emph{cardinality constraint} is a rule of the following form:
\begin{equation} 
 ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u \leftarrow body.
 \label{eq7ASP}
\end{equation}
with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$). \repEB{ This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$. But that the number of this atoms $q_i$ should be between $l$ and $m$. That means that it is possible to find the same atom repeated many times at one answer set.}  \\
In addition the literals $q_i$ can also appear in the body of the rule,
and $l$ and $u$ are the lower and upper bounds of the number of these literals
that can appear in the answer sets.
%cardinality of the subsets of $ \{q_{1}, ... , q_{m}\}$ that satisfy the answer sets.
Therefore, a cardinality constraint is satisfied in an answer set $X$, if the number of atoms of $\{q_{1}, ... , q_{m}\}$ belonging to $X$ is between $l$ and $u$,
or, in other words, if:
\begin{equation}
  l \leq | \{q_{1},... , q_{m}\} \cap X | \leq u
\end{equation}
where $\cap$ is the symbol of the intersection of sets
and $|A|$ is the cardinality of the set $A$.
\modMF{%
In the following, if they are not explicitly given,
we consider that $l$ defaults to $0$ and $u$ defaults to $\infty$.
\annotMF{Les cardinalités ne sont pas seulement utilisées de cette manière ! Dans notre code, on les trouve dans la tête et dans le corps des règles.}
\todo{À compléter : cardinalités dans le corps des règles}
}%
\end{itemize}

@@

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}:
\begin{itemize}
\item it is a simplification of a given system,
\item it allows an action on the system \repEB{like resolving problems, searching for a path, identifying states ... }
\annotMF{Quel genre d'action ?} 
\end{itemize}
This concept model can address another angle issues related to representation process.

@@

\item \textbf{Solving an ASP program}
\modMF{%
The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
that usually work together:
}%
first the grounder is used to remove variables in order to achieve an equivalent but constant program,
then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
Amongst all available grounders for ASP, we can name
\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
and the solvers include
\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...

\reformuler{
The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations. Both systems employ grounding algorithms that work quickly and simplify the program.
}
\repEB{c'est à ajouter dans le paragraphe ou à reformuler?}
For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

>>>>

This simple language has many advantages that make it very effective in terms of knowledge representation, reasoning and declarative problem solving.

@@

\annotMF{Il faut absolument décider d'une charte d'écriture pour les programmes ASP dans ce texte, d'autant plus qu'actuellement cette charte change entre la section courant (où le code est écrit en mode \$maths\$) et la suite où il est en mode \texttt{listing}.}
\repEB{à revoir}
The other classes remain constant as they meet some informal agreement. In general, variables start with a capital letter and contain letters and numbers, but often only one capital letter is used (\eg X, Y, ...).
The constants, symbols and predicates follow the same rule, but they begin with a lowercase letter (\eg f, g, a, b, ...).
String surrounded by quotation marks are also considered as constants (\eg "f", "g", "a b c", ...).

@@

\item A \emph{cardinality constraint} is a rule of the following form:
\begin{equation} 
 ~ l ~\{q_{1},~ ... ,~ q_{m}\}~ u \leftarrow body.
 \label{eq7ASP}
\end{equation}
with $m \geq 1$, $l$ an integer and $u$ an integer or the infinity ($\infty$). This form of rule \ref{eq7ASP} is used when the body can lead to several $q_i$ atoms and also when the programmer need to have more than an atom of type $q_i$. But that the number of this atoms $q_i$ should be between $l$ and $m$. That means that it is possible to find the same atom repeated many times at one answer set.\\
In addition the literals $q_i$ can also appear in the body of the rule,
and $l$ and $u$ are the lower and upper bounds of the number of these literals
that can appear in the answer sets.
%cardinality of the subsets of $ \{q_{1}, ... , q_{m}\}$ that satisfy the answer sets.
Therefore, a cardinality constraint is satisfied in an answer set $X$, if the number of atoms of $\{q_{1}, ... , q_{m}\}$ belonging to $X$ is between $l$ and $u$,
or, in other words, if:
\begin{equation}
  l \leq | \{q_{1},... , q_{m}\} \cap X | \leq u
\end{equation}
where $\cap$ is the symbol of the intersection of sets
and $|A|$ is the cardinality of the set $A$.
In the following, if they are not explicitly given,
we consider that $l$ defaults to $0$ and $u$ defaults to $\infty$.

\end{itemize}

@@

We can consider that the construction of models is one of the fundamental components of the scientific process. It concerns all systems we seek to control. A model has two main features \cite{Glimpse}:
\begin{itemize}
\item it is a simplification of a given system,
\item it allows an action on the system like resolving problems, searching for a path, identifying states and verufying properties... 
\end{itemize}
This concept model can address another angle issues related to representation process.

@@

\item \textbf{Solving an ASP program}
The solving of ASP programs relies on a \emph{grounder} and a \emph{solver}
that usually work together:
first the grounder is used to remove variables in order to achieve an equivalent but constant program,
then the solver computes all answer sets for a stabilized program generated by the grounder~\cite{Vladimir,AnsPrologAPE}.
Amongst all available grounders for ASP, we can name
\texttt{Gringo}, \texttt{DLV} and \texttt{LParse},
and the solvers include
\texttt{SModels}, \texttt{DLV}, \texttt{CModels}, \texttt{Clasp}...

\reformuler{
The combined use of the grounder and the solver specifies major programs in a compact, using rules with schematic variables and other abbreviations. Both systems employ grounding algorithms that work quickly and simplify the program.
}
For the work presented in this paper, we worked with \texttt{Clingo} which is a combination of the grounder \texttt{Gringo} and the solver \texttt{Clasp}.

### 2015-01-26 PH.tex
update

\pref{def:PH} introduces the Process Hitting~(PH)~\cite{PMR10-TCSB}
which allows to model
%A Process Hitting (PH) (\pref{def:PH}) gathers 
a finite number of local levels,
called \emph{processes},
grouped into a finite set of components, called \emph{sorts}.
A process is noted $a_i$, where $a$ is the sort's name,
and $i$ is the process identifier within sort $a$.
At any time, exactly one process of each sort is \emph{active},
and the set of active processes is called a \emph{state}.

The concurrent interactions between processes are defined by a set of \emph{actions}.
Actions describe the replacement of a process by another of the same sort
conditioned by the presence of at most one other process in the current state.
An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
“$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$”,
where $a_i$, $b_j$, $b_k$ are processes of sorts $a$ and $b$,
called respectively \emph{hitter}, \emph{target} and
\emph{bounce} of the action.
We also call a \emph{self-hit} any action whose hitter and target sorts are the same,
that is, of the form: $\PHfrappe{a_i}{a_i}{a_k}$.

@@

\path[bounce,bend left]
\TBounce{a_0}{}{a_1}{.south}
;
\TState{a_0,b_0,z_1}
\end{tikzpicture}
\caption{\label{fig:ph} 
A PH model example with three sorts: $a$, $b$ and $z$ ($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2). Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 5 \emph{actions} that model the dynamic behavior are depicted by pairs of arrows in solid and dotted lines. The grayed processes stand for the possible initial state: $\PHstate{a_1, v_0, z_1}$.
}
\end{figure}
\end{example*}
A state of the networks is a set of active processes containing a single process of each sort.
\modMF{%
The active process of a given sort $a \in \PHs$ in a state $s \in \PHl$
is noted $\PHget{s}{a}$.
For any given process $a_i$ we also note: $a_i \in s$ if and only if $\PHget{s}{a} = a_i$.
}%

@@

\begin{definition}[Next states]
\label{def:NextState}
Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
one of its states. The set of the next possible states for $s$ are computed as follows:
\begin{center}
$next(s) = \{s[b_k /b_j ]| \exists (a_i , b_j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa\}$
\end{center}
\todo{Définir la notation $s[b_k /b_j ]$}
\annotMF{Est-ce que $next(s)$ est vraiment utile ? On ne s'en ressert pas après.}
\end{definition}

@@

\begin{definition}[Stable state or fix point]
\label{def:FixPoint1}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be state;
$s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
\annotMF{“Stable state” ? “Fix point” ? “Fixed point” ?}
\end{definition}

@@

\begin{definition}[Reachability question]
\label{def:reachability}

 If $s \in \PHl$ is a state and $A \subseteq \Proc$ a set of processes,
 we note $\mathcal{P}(s, A)$ the \emph{the reachability question}:
 \begin{center}
  $\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$
 \end{center}
\todo{Il faut définir $\Proc$}
\annotMF{Il serait peut-être mieux de parler de “reachability property” plutôt que “question”.}
\end{definition}

>>>>

\pref{def:PH} introduces the Process Hitting~(PH)~\cite{PMR10-TCSB}
which allows to model a finite number of local levels,
called \emph{processes},
grouped into a finite set of components, called \emph{sorts}.
A process is noted $a_i$, where $a$ is the sort's name,
and $i$ is the process identifier within sort $a$.
At any time, exactly one process of each sort is \emph{active},
and the set of active processes is called a \emph{state}.

The concurrent interactions between processes are defined by a set of \emph{actions}.
Each action is responsible for the replacement of one process by another of the same sort
conditioned by the presence of at most one other process in the current state.
An action is denoted by $\PHfrappe{a_i}{b_j}{b_k}$, which is read as
“$a_i$ \emph{hits} $b_j$ to make it \emph{bounce} to $b_k$”,
where $a_i$, $b_j$, $b_k$ are processes of sorts $a$ and $b$,
called respectively \emph{hitter}, \emph{target} and
\emph{bounce} of the action.
We also call a \emph{self-hit} any action whose hitter and target sorts are the same,
that is, of the form: $\PHfrappe{a_i}{a_i}{a_k}$.

@@

\path[bounce,bend left]
\TBounce{a_0}{}{a_1}{.south}
;
\TState{a_0,b_0,z_1}
\end{tikzpicture}
\caption{\label{fig:ph} 
A PH model example with three sorts: $a$, $b$ and $z$ ($a$ is either at level 0 or 1, $b$ at either level 0 or 1 and $z$ at either level 0, 1 or 2). Boxes represent the \emph{sorts} (network components), circles represent the \emph{processes} (component levels), and the 5 \emph{actions} that model the dynamic behavior are depicted by pairs of arrows in solid and dotted lines. The grayed processes stand for the possible initial state: $\PHstate{a_1, v_0, z_1}$.
}
\end{figure}
\end{example*}
A state of the networks is a set of active processes containing a single process of each sort.
The active process of a given sort $a \in \PHs$ in a state $s \in \PHl$
is noted $\PHget{s}{a}$.
For any given process $a_i$ we also note: $a_i \in s$ if and only if $\PHget{s}{a} = a_i$.

@@

\begin{definition}[Next states]
\label{def:NextState}
Let $(\PHs,\PHl,\PHa)$ be a Process Hitting and $s \in \PHl$ be
one of its states. The set of the next possible states for $s$ are computed as follows:
\begin{center}
$next(s) = \{s[b_k /b_j ]| \exists (a_i , b_j ) \in s^2 , \exists b_k \in \PHl_b , \PHfrappe{a_i}{b_j}{b_k} \in \PHa\}$
\end{center}
\todo{Définir la notation $s[b_k /b_j ]$}
\annotMF{Est-ce que $next(s)$ est vraiment utile ? On ne s'en ressert pas après.} \repEB{Je crois que si, pour le définition du point fixe et pour comprendre l'évolution du réseau d'un état à un autre}
\end{definition}

@@

\begin{definition}[Stable state or fixed point]
\label{def:FixPoint1}
Let $PH = (\PHs,\PHl,\PHa)$ be a Process Hitting and
$s \in \PHl$ be state;
$s$ is a \emph{stable state} for $PH$ if and only if $next(s) = \emptyset $.
\end{definition}

@@

\begin{definition}[Reachability property]
\label{def:reachability}

 If $s \in \PHl$ is a state and $A \subseteq \Proc$ with $\Proc$ is a set of processes that we find at most one process of each sort $\{a_i, b_j,..., z_k\}$ 
 we note $\mathcal{P}(s, A)$ the \emph{the reachability property}:
 \begin{center}
  $\exists? \delta \in \Sce(s), \forall a_i \in A, \PHget{(s \play \delta)}{a}=a_i$
 \end{center}
\todo{Il faut définir $\Proc$}
\end{definition}

### 2015-01-26 comparaison.tex
update tableau et abstract
update

In this section, we show the effectiveness of our approach on some examples,
\modMF{%
and compare it to other existing approaches.
}%
All computations were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al. \cite{roccaasp} which was developed ALSO on ASP.
\annotMF{Il y a une confusion entre méthodes et outils ici je pense.} \repEB{J'ai modifié}
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
Each method use a specific kind of representation:
logical regulatory network for GINsim,
Instantiable Transition Systems for LIBDDD,
state transition network for Rocca's et al. method
and Process Hitting (PH) for Pint as well as for our method.

\modMF{%
For this comparative study, we focus biological network of different sizes:
a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
an ERBB receptor-regulated G1/S transition (ERBB) model with 20 components~\cite{Samaga2009}
and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.
}%
These models were chosen to be of different sizes:
from small (12 components) to large (40 components).
We note however that the PH models considered may contain more sorts than
the original number of biological components, due to the use of
“cooperative sorts”, which allow to model Booelan gates but do not necessarily
have a biological meaning.
The different model representations required have been obtained by translations
from the PH
ensuring the conservation of the dynamical properties.
All these translations are available with \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} }
\todo{URL vers Pint et les autres outils}
All results alongside with more detailed specifications of the models
are given in table \ref{tab:reachability}.
The methods and the results provided by each of them are detailed in the following.
The overall results shows that our method is efficient in computing reachability
from a given initial state;
furthermore, it sometimes provides more informations that the other existing ones.

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & \todo{XX} & \todo{XX} & 0m01.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m01.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m05.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Rajouter le nombre de sortes dans le tableau ou la description} \repEB{avec ou sans les sortes coopératives? car pour PH y'en a plus de composants que pour les autres modèles}
\annotMF{Il faut expliciter ce qu'on entend par “state” et “process”. Il fudrait peut-être aussi changer les termes ; je propose : “full state” et “partial state”.} \modifEB{Je préfère ta proposition mais tu vois ça prend plus d'espace et le tableau dépasse les limites }

@@

\item \textbf{Rocca's method} offers the possibility to model-check CTL properties
of state-transition networks.
\modMF{%
We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
Therefore, using this method first requires an exponential computation to
express the state-transition graph of the studied model into ASP.
Indeed, for the tadpole tail resorption model comprising only 42 processes,
this computation took 3 minutes, which was not included in the solving time
reported in table~\ref{tab:reachability}.
Afterwards, this method is able to check any kind of CTL formula
(and not only the EF form that we focused on in this paper),
but this checking may be time-consuming.
}%

@@

\item \textbf{\textsc{libddd}}
\modMF{%
is a library for symbolic model-checking of CTL \& LTL properties.
It can thus especially be used to check reachability properties;
however, as opposed to our method, it does not output an execution path
solving this reachability.
In addition, it relies on the construction of the state-transition graph
which is then stored under the form of a binary decision diagram for a more efficient analysis.
}%
This computation explains why \textsc{libddd} takes more time to respond,
and does not answer in less than 12 minutes for the biggest example
(which contains $2^{73}$ states).
Finally, \textsc{libddd} is not able to compute the stable states of a network.

\item \textbf{PINT}
\modMF{%
is a library gathering tools and translators related to the PH.
It should be noted that the only reachability analysis developed so far on the Process Hitting
was implemented in the software Pint, and consists in an approximation:
by avoiding to compute the state-transition graph, it is ensured to be really efficient,
at the cost of possibly terminating without being conclusive (although this is rare).
}%
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
In the first example \textsc{PINT} did not give an answer because of the number of goals is greater than 5.
\end{itemize}

@@

We developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
\modMF{%
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
}%
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop. This is possible with the option \texttt{"--imax=n"}.
\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
\annotMF{Cette ligne de commande ne me semble pas nécessaire. Mais si on la garde, il faut en donner un autre exemple plus tôt, sans l'argument -{}-imax=n.} \modifEB{ok t'as raison je l'ai enlevée et remplacée par juste l'option \texttt{"--imax=n"}}
Some interesting values can be given to this integer \texttt{n},
which sets the maximum number of steps.
For example, the total number of states is a maximum which is obvious,
as it will never be exceeded by a minimum path,
but too hight to be very interesting.
The total number of sorts can also be an interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks,
or, by using the same reasoning, the total number of processes.
\modMF{%
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
}%

>>>>

In this section, we show the effectiveness of our approach on some examples,
and compare it to other existing approaches.
All computations were performed on a Pentium~V, 3.2~GHz with 4~GB RAM.

\subsection{Evaluation}
To evaluate the effectiveness of our new approach,
we position ourselves with respect to existing methods dealing with different biological models.
We have chosen the following tools, that are detailed below: 
\textsc{GINsim} (Gene Interaction Network Simulation)~\cite{gonzalez2006ginsim,naldi2009logical,naldi2007decision},
\texttt{LIBDDD} (Library of Data Decision Diagrams)~\cite{thierry2009hierarchical,colange2013towards},  PINT \cite{PMR12-MSCS} and the method of Rocca's et al. \cite{roccaasp} which was developed also on ASP.
\annotMF{Il faut trouver un autre nom pour “Rocca's method”.} \repEB{c'est vrai ; "transition state networks with ASP?}
Each method use a specific kind of representation\footnote{The transalation from other models to PH is done through the converters availbale on line: \url{https://github.com/pauleve/pint/tree/master/converters}}:
logical regulatory network for \textsc{GINsim} \footnote{\textsc{GINsim} is a computer tool for the modeling and simulation of genetic regulatory networks \url{http://ginsim.org/} },
Instantiable Transition Systems for \textsc{libDDD} is a library for symbolic model-checking of CTL \& LTL properties. \footnote{\textsc{libDDD} is \url{http://www.lip6.fr/} },
state transition network for Rocca's et al. method
and Process Hitting (PH) for \textsc{PINT} \footnote{\textsc{PINT} is a Process Hitting tool that implements the language, simulation, formal analysis, and translation of PH models. It available online \url{http://loicpauleve.name/pint/} } as well as for our method.

For this comparative study, we focus biological network of different sizes:
a tadpole tail resorption (TTR) model with 12 biological components~\cite{khalis2009smbionet},
an ERBB receptor-regulated G1/S transition (ERBB) model with 20 components~\cite{Samaga2009}
and a T-cell receptor (TCR) signaling network of 40 components~\cite{Klamt06}.

These models were chosen to be of different sizes:
from small (12 components) to large (40 components).
We note however that the PH models considered may contain more sorts than
the original number of biological components, due to the use of
“cooperative sorts”, which allow to model Booelan gates but do not necessarily
have a biological meaning.
The different model representations required have been obtained by translations
from the PH
ensuring the conservation of the dynamical properties.
All these translations are available with \textsc{PINT} 
All results alongside with more detailed specifications of the models
are given in table \ref{tab:reachability}.
The methods and the results provided by each of them are detailed in the following.
The overall results shows that our method is efficient in computing reachability
from a given initial state;
furthermore, it sometimes provides more informations that the other existing ones.

\begin{figure}[htp]
\begin{center}
\label{tab:reachability}
\noindent%
\begin{tabular}{|l|c|c|c|c||c|c|c|c|c|}
  \hline
   Model&  \#sorts &\#procs & \#states & target & Rocca & Pint & libddd & GINsim & ASP \\
  \hline
  TTR & 12 &42 & $2^{19}$ & full state & 7m17.01s & - & 0m1.151s &  0m1.001s & 0m1.90s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & full state & - & - &1m55.38s & 2m31.64s & 0m11.84s \\
  \hline
  ERBB & 20 &152 & $2^{70}$ & partial state & - & 0m0.027s &1m54.96s & - & 0m5.02s \\
  \hline
  TCR & 40 &156 & $2^{73}$ & full state & - & - & out & $ \lbrack 5mn ; \infty \rbrack$ & 4m27.93s \\
  \hline
   TCR & 40 &156 & $2^{73}$ & partial state & - & 0m0.014s & out & - & 1m35.080s \\
  \hline
\end{tabular}
\caption{Compared performances of Rocca et al. method, \textsc{Pint}, \textsc{libddd}, \textsc{GINsim} and our new method developped in ASP.
For each test, this table gives the considered model,
the number of its processes and its states, the type of goal
(either a whole state or a sub-state)
and the computation time of the different methods used for the tests,
where “out” marks an execution taking too much time or memory
and “-” indicates that is not possible to do the test.
}
\end{center}
\end{figure}

\todo{Diviser en 2 tableaux}

@@

\item \textbf{Rocca's method} offers the possibility to model-check CTL properties
of state-transition networks.
We note that in general, computing such a graph from a network is exponential in the number of components and local states in the model.
Therefore, using this method first requires an exponential computation to
express the state-transition graph of the studied model into ASP.
Indeed, for the tadpole tail resorption model comprising only 42 processes,
this computation took 3 minutes, which was not included in the solving time
reported in table~\ref{tab:reachability}.
Afterwards, this method is able to check any kind of CTL formula
(and not only the EF form that we focused on in this paper),
but this checking may be time-consuming.

@@

\item \textbf{\textsc{libddd}}
is a library for symbolic model-checking of CTL \& LTL properties.
It can thus especially be used to check reachability properties;
however, as opposed to our method, it does not output an execution path
solving this reachability.
In addition, it relies on the construction of the state-transition graph
which is then stored under the form of a binary decision diagram for a more efficient analysis.

This computation explains why \textsc{libddd} takes more time to respond,
and does not answer in less than 12 minutes for the biggest example
(which contains $2^{73}$ states).
Finally, \textsc{libddd} is not able to compute the stable states of a network.

\item \textbf{PINT}
is a library gathering tools and translators related to the PH.
It should be noted that the only reachability analysis developed so far on the Process Hitting
was implemented in the software Pint, and consists in an approximation:
by avoiding to compute the state-transition graph, it is ensured to be really efficient,
at the cost of possibly terminating without being conclusive (although this is rare).
Moreover, in the case of a positive answer,
it currently does not return the execution the path achieving the desired reachability,
but only outputs its conclusion.
In the first example \textsc{PINT} did not give an answer because of the number of goals is greater than 5.
\end{itemize}

@@

We developed new methods to check dynamical properties,
namely identifying stable states and finding all possible paths to reach a given goal.
Compared to other methods (Rocca's method, \textsc{GINsim} and \textsc{libddd}),
our method is relatively faster and also permits to study larger networks
(up to $2^{73}$ states in our tests) despite some inconclusive cases.
Indeed, our ASP program is non-conclusive in the cases where the given goal is not
reachable and the model contains loops in its dynamics
(which happens in almost every biological model).
In this case, the program will compute all infinite paths in these loops,
and never reach a goal or a fixed point.
It is still possible, however, to limit the number of iterations to an arbitrary
maximum which will be eventually reached in the case of an endless loop. This is possible with the option \texttt{"--imax=n"}.
\annotMF{On peut éventuellement citer quelques approximations justifiées de ce maximum.}
%%
%\begin{lstlisting}[numbers=none]
%iclingo find\_path.lp <network\_name>.lp -- imax=n
%\end{lstlisting}
%%
Some interesting values can be given to this integer \texttt{n},
which sets the maximum number of steps.
For example, the total number of states is a maximum which is obvious,
as it will never be exceeded by a minimum path,
but too hight to be very interesting.
The total number of sorts can also be an interesting value,
under the hypothesis that each one will change its active process at most once,
which is often the case for Boolean networks,
or, by using the same reasoning, the total number of processes.
In this case, however, a termination with no solution cannot be considered as a formal
negative answer, unless one can prove that the chosen value \texttt{n}
is bigger than the longest possible path in the state-transition graph.
Given our implementation, if the step \texttt{n} is reached,
the computation stops and displays paths that are not valid executions
to the goal.
